BarnebyLives: an R package to create herbarium specimen labels and clean spreadsheets

### 2025-04-25 manuscript/BarnebyLives An R package to create and qc herbarium labels.rmd
part of letter to editor; v difficult to not be cheeky given why this actually exists...
letter to editor
formatting

subs_batch <- c(11207,  1263, 1326, 1384, 1388, 1394, 1396, 1397, 1400, 1435, 1437, 1438,
                2161, 2301, 2517, 2746, 2785)

@@

LaTeX offers well-documented and detailed functionality to customize labels for individuals and institutions. 
Anecdotally, using its default settings, it can produce more aesthetically pleasing results than typical word processors. 
Very good documentation of LaTeX capabilities is offered in multiple areas; for instance, via the [Overleaf](https://www.overleaf.com/learn) project.

# CONCLUSIONS
BarnebyLives is an R package that can be used to rapidly acquire relevant geographic and taxonomic data. 
It can also perform specialized spell checks and assorted curatorial tasks to produce both digital and analog data. 
The package relies on no licensed software, such as the Microsoft Office suite, and is suitable for install on all major operating systems  (Windows, Mac, Linux), however currently label generation support is only offered on Linux and Mac, with a small amount of use of the command line, which may be called from the Rstudio rather than a 'traditional' terminal.

@@

# ACKNOWLEDGEMENTS 
The Bureau of Land Management is gratefully acknowledged as a provider of funding to R. C. B. for most of his specimen collection activities. 
We thank the two anonymous peer reviewers who have increased the quality of this manuscript.
Several prominent associated collectors of specimens used in this study are thanked: Dani Yashinovitz, Hannah Lovell, Dakota Becerra, Caitlin Miller, Hubert Szczygiel. 
We thank Rosalind Rowe for providing useful feedback during the later developmental stages of the program.

>>>>

subs_batch <- c(1586, 1880, 2187, 2159)

@@

![A label generated from a default template](../graphics/plots/RCB2759.png){width=50%}

LaTeX offers well-documented and detailed functionality to customize labels for individuals and institutions. 
Anecdotally, using its default settings, it can produce aesthetically pleasing results (Figure 4). 
Very good documentation of LaTeX capabilities is offered in multiple areas; for instance, via the [Overleaf](https://www.overleaf.com/learn) project.

# CONCLUSIONS
BarnebyLives is an R package that can be used to rapidly acquire relevant geographic and taxonomic data. 
It can also perform specialized spell checks and assorted curatorial tasks to produce both digital and analog data. 
The package relies on no licensed software, such as the Microsoft Office suite, and is suitable for install on all major operating systems  (Windows, Mac, Linux), however currently label generation support is only offered on Linux and Mac, with a small amount of use of the command line, which may be called from the Rstudio rather than a 'traditional' terminal. 
Thematically, it's area of interest is Western North America.

@@

# ACKNOWLEDGEMENTS 
The Bureau of Land Management is gratefully acknowledged as a provider of funding to R.C.B. for most of his specimen collection activities. 
We thank the two anonymous peer reviewers who have increased the quality of this manuscript.
Several prominent associated collectors of specimens used in this study are thanked: Dani Yashinovitz, Hannah Lovell, Dakota Becerra, Caitlin Miller, Hubert Szczygiel.

### 2025-04-25 manuscript/CoverLetter.rmd
formatting

While a variety of software tools exist for interacting with accessioned herbarium data, few are dedicated to accessioning new material. 
And even fewer of these tools are intended for an actual collector to use.  
This isn't surprising, as the number of collectors who are not curators has decreased, and the barrier to entry for both activities is a degree of particularity and eccentricity. 
Unfortunately, the particularity of curators often makes contributing to herbaria a challenging task for collectors. 
A common challenge for collectors, especially those who work on large projects implemented by federal agencies, such as germplasm collection (e.g. SOS) and ecological monitoring (e.g. NWCA, and AIM), is the need to simultaneously satisfy the particularities of multiple curators under tight deadlines.

Fortunately, I have been scolded by many curators over the last decade, on occasion even receiving constructive criticism, and have built up a wide knowledge of quirks to cater to. 
To assist in running quality control checks, for both my - and the young botanists whom I manage - collections, I developed a wide range of functions which help meet a broad set of interests.
In fact 'BarnebyLives' the package which integrates these tools, has 30 user facing functions,  10+ internal functions, and 8 built in data sets.

@@

\newpage

Reviewers:

Shelly Gaynor mlgaynor@umich.edu 
R package author extraordinaire, including multiple documented in APPS software notes, and (former?) member of iDigBio team.

Ben Legler blegler@uwyo.edu 
Database administrator and creator of two large herbaria consortia websites, and a *bona fide* field botanist.

Jason Alexander jason_alexander@berkeley.edu 
I don't actually know Jason, word from our mutuals is he knows his stuff and this should be within his wheelhouse.

Charles Davies cdavis@oeb.harvard.edu 
A bit of a stretch, but I need to suggest four reviewers. 
I doubt he has time for this, but maybe one of his graduate students would be interested. 
Reasoning: Charles is a profuse user of herbarium specimens for both global scale ecology studies, localized natural history, and systematics.

>>>>

While a variety of software tools exist for interacting with accessioned herbarium data, few are dedicated to accessioning new material. 
And even fewer of these tools are intended for an actual collector to use. 
Which isn't surprising, as the number of collectors who are not curators has decreased, and the barrier to entry for both activities is a high degree of particularity and eccentricity. 
Unfortunately, the particularity of curators often makes contributing to herbaria a challenging task for collectors. 
A common challenge for collectors, especially those who work on large projects implemented by federal agencies, such as germplasm collection (e.g. SOS) and ecological monitoring (e.g. NWCA, and AIM), is the need to simultaneously satisfy the particularities of multiple curators under tight deadlines.

Fortunately, I have been scolded by many curators over the last decade, and on occasion received actionable criticism, and have built up a wide knowledge of quirks to cater to. 
To assist in running quality control checks, for both my - and the young botanists whom I manage - collections, I developed a wide range of functions which help meet a broad set of interests.
In fact 'BarnebyLives' the package which integrates these tools, has 30 user facing functions,  10+ internal functions, and 8 built in data sets.

@@

Sincerely,  
Reed
